---
title: "Simulate pupil data"
author: "Rich"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As the title suggests.
This file simulates some rough and ready pupil data for one eye and then for both eyes.
The purpose of simulating data is to test a preporcessing pipeline, hence it does not need to be overly realistic of real pupil size data.

# 1. load packages and custom functions plus adjust settings #

## packages ##

```{r load-pkg, results=F, message=F, warning=F}
pkg <- c("tidyverse", "here", "future", "RColorBrewer", "parallel", "patchwork") 

lapply(pkg, library, character.only = TRUE)
```

## custom functions ##

```{r functions}
source(here("custom_functions", "simulate.r"))
source(here("custom_functions", "math.r"))
source(here("custom_functions", "settings.r"))
```

## check multicore settings ##

```{r settings}
plan(multicore)
message(paste("Supporting multicore:", supportsMulticore()))
message(paste("Available cores:", detectCores()))
```

# 2. simulate pupil data for one eye #

```{r sim-pupil-data}
sim_pupil_data <- generate_pupil_data(n_participants = 3, n_trials = 3, hz = 60,
                                      eye = "left", trial_length = 120, seed = 123)
head(sim_pupil_data)
# glimpse(sim_pupil_data)
```

summarise

```{r summarise}
# Preview the data summary
summary <- sim_pupil_data %>%
  group_by(participant_id, trial_id) %>%
  summarize(
    n_samples = n(),
    min_time = min(time_ms),
    max_time = max(time_ms),
    n_blinks = sum(is.na(pupil_size)),  # Count missing values (blinks)
    .groups = "drop"
  )
summary
```

take a look and plot artifacts

```{r plot, echo=F}
# Plot example participant's trials with artifacts
p2.1 <- ggplot(sim_pupil_data,
               aes(x = time_ms, y = pupil_size, color = trial_id)) +
  geom_line() +
  geom_point(data = . %>% filter(is.na(pupil_size)), 
             aes(x = time_ms, y = 3), color = "red", size = 2) +  # Mark blinks in red
  labs(title = "Simulated Pupil Size Data (60Hz) with Blinks & Artifacts", 
       x = "Time (ms)", y = "Pupil Size (mm)") +
  scale_color_brewer(palette = "Dark2") +
  facet_grid(participant_id~trial_id)
p2.1

ggsave("figures/sim_pupil_data.png",
       width = 10, height = 7)
```

add a message column and wrangle a little

```{r wrangle}
sim_pupil_data <- sim_pupil_data %>% 
  mutate(message = case_when(
    time_ms < 1000 ~ "fixation",
    .default = "target"),
    time_ms = custom_round(time_ms)) %>% 
  rename(subject = participant_id, trial = trial_id, time = time_ms, 
         pupil = pupil_size) %>% 
  ## change subject and trial to as.double and values
  mutate(subject = parse_number(subject),
         trial = parse_number(trial)) 
head(sim_pupil_data)
```

write out a file

```{r write-out}
write_csv(sim_pupil_data, "data/sim/sim_pupil_data.csv")
```

# 3. simulate pupil data for both eyes #

```{r sim-pupil-data-both}
sim_pupil_data_both <- generate_pupil_data(
  n_participants = 3,
  n_trials = 3, 
  eye = "both",       # Simulate both eyes
  hz = 60,            # Sampling rate
  trial_length = 120, # Each trial is 120 samples (2 seconds at 60Hz)
  seed = 123          # make it reproducible         
)
head(sim_pupil_data_both)
```

summarise

```{r summarise-both}
# Preview the data summary
summary_both <- sim_pupil_data_both %>%
  group_by(participant_id, trial_id, eye) %>%
  summarize(
    n_samples = n(),
    min_time = min(time_ms),
    max_time = max(time_ms),
    n_blinks = sum(is.na(pupil_size)),  # Count missing values (blinks)
    .groups = "drop"
  )
summary_both
```

take a look and plot artifacts

```{r plot-both, echo=F}
# Plot example participant's trials with artifacts
p3.1 <- ggplot(sim_pupil_data_both,
               aes(x = time_ms, y = pupil_size, color = eye)) +
  geom_line() +
  geom_point(data = . %>% filter(is.na(pupil_size)), 
             aes(x = time_ms, y = 3), color = "red", size = 2) +  # Mark blinks in red
  labs(title = "Simulated Pupil Size Data (60Hz) with Blinks & Artifacts", 
       x = "Time (ms)", y = "Pupil Size (mm)") +
  scale_color_brewer(palette = "Dark2") +
  facet_grid(participant_id~trial_id)
p3.1

ggsave("figures/sim_pupil_data_both.png",
       width = 10, height = 7)
```

add a message column and wrangle a little

```{r wrangle-both}
sim_pupil_data_both <- sim_pupil_data_both %>% 
  mutate(message = case_when(
    time_ms < 1000 ~ "fixation",
    .default = "target"),
    time_ms = custom_round(time_ms)) %>% 
  rename(subject = participant_id, trial = trial_id, time = time_ms, 
         pupil = pupil_size) %>% 
  ## change subject and trial to as.double and values
  mutate(subject = parse_number(subject),
         trial = parse_number(trial)) %>% 
  pivot_wider(
    id_cols = c(subject, trial, time, message),
    names_from = "eye",
    values_from = c("x_pos", "y_pos", "pupil")
  )
head(sim_pupil_data_both)
# glimpse(sim_pupil_data_both)
```

write out a file

```{r write-out-both}
write_csv(sim_pupil_data_both, "data/sim/sim_pupil_data_both.csv")
```
